%\documentclass[useAMS,usenatbib,preprint2]{aastex}
\documentclass{emulateapj}
% \documentclass[12pt,preprint]{aastex}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{times}
\usepackage{natbib}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsbsy}

\bibliographystyle{apj}

%%%%%%%%%%%%%%%%

\begin{document}
\title{Intermittency of the 625 Hz QPO in the 2004 Hyperflare by SGR 1806-20}
\author{Daniela Huppenkothen}
\affil{Instituut Anton Pannekoek, University of Amsterdam, Amsterdam  1098 XH,
  The Netherlands}
\email{d.huppenkothen@@uva.nl}
\author{Anna Watts}
\affil{Instituut Anton Pannekoek, University of Amsterdam, Amsterdam  1098 XH,
  The Netherlands}
\author{Yuri Levin}
\affil{Monash}

\begin{abstract}
ABSTRACT GOES HERE!
\end{abstract} 

\keywords{pulsars: individual (SGR 1806--20), stars:
  magnetars, stars: oscillations, X-rays: stars}
\section{Introduction}
\label{sec:introduction}


\section{Data Analysis}
\label{sec:analysis}

We include data sets from two different space telescopes in our analysis: The {\it Rossi} X-ray timing Explorer (RXTE), and the {\it Ramaty High Energy Solar Spectroscopic Imager (RHESSI)}. An overview of the RXTE data is given in \citep{israel05}. Data was recorded in $\mathrm{Goodxenon_2s mode}$, allowing for time resolution up to $1 \, \mu \mathrm{s}$, high enough to study high-frequency QPOs.
Observations taken with {\it RHESSI} are detailed in \citep{watts06}. Following their analysis, we only used photons recorded with the eight front segments of the telescope, which are largely uncontaminated by scattering in the back of the space craft. The high-frequency QPO is seen only in the energy range between $100 \, \mathrm{keV}$ and $200 \, \mathrm{keV}$, hence we filter for these energies. All data are barycentered, that is, corrected for the motion of the space craft through space to avoid systematic effects in the timing analysis.

For the RXTE data, we concentrated on the part of the light curve where the $625 \, \mathrm{Hz}$ QPO was originally found, from around $190\, \mathrm{s}$ after the onset of the flare to the end of the observation. This encompasses a total of $15$ rotational cycles of the neutron star. The {\it RHESSI} observations place the same QPO at a slightly different frequency ($626.5 \, \mathrm{Hz}$ as opposed to $625.5 \, \mathrm{Hz}$), and at an earlier time. For the latter, we search the range from $80\, \mathrm{s}$ to $225 \, \mathrm{s}$ from the onset of the flare, or equivalently 19 cycles.

We split each rotational cycle into a number $N_\mathrm{r}$ of overlapping segments of length $\delta t_{\mathrm{s}}$. For each of these segments, we computed the periodogram and extracted the power at the frequency where the QPO was observed. For each periodogram we tested the significance of the observed power against $N_{\mathrm{sim}}$ simulations, which are constructed in the following way.

As a first step, we smoothed out the light curve to a resolution of $0.01 \, \mathrm{s}$, or equivalently $100 \, \mathrm{Hz}$, ensuring that all possible variability at smaller time scales is eliminated from the data. We then interpolated back to the original time resolution used ($\delta t = 2.5 \times 10^{-5} \, \mathrm{s}$), and added Poisson noise to this smoothed light curve $N_{\mathrm{sim}}$ times. This represents the null hypothesis that the QPO is not present, and that any variability measured at $625 \, \mathrm{Hz}$ is solely due to photon counting noise in the detector. For each of our $N_{\mathrm{sim}}$ simulations, we performed exactly the same analysis as for the observed data. We can then compare the real powers we measured for a given segment to the distribution of simulated powers in that segment. Additionally, we can compare the maximum power observed at $625 \, \mathrm{Hz}$ for all segments in our observed data for the maximum powers at this frequency in the ensemble of simulated light curves. This allows us to construct a p-value for the significance of the observed maximum power to be consistent with the null hypothesis, i.e. pure noise. Note that a number of trials correction for the number of segments searches is automatically included, since we apply the same search procedure to the simulations.


In addition to testing all segments individually, we constructed phase-averaged periodograms in the following way. We match all periodograms belonging to segments that start at the same phase with each other. In order to construct the two-cycle average, we average the same phase bins for two consecutive cycles together, and again extract the power at the relevant frequency. We then do the same for the next cycle, and so on until we reach the end of the data under consideration. The result is a moving average over subsequent rotational cycles, where the averaged periodograms match in phase.
Similarly, we can construct three-cycle averages by combining peridoograms from three consecutive cycles, and so forth, until we average the maximum number of cycles in our particular data set. Note that in both cases, the observed (averaged) powers are not independent. The powers extracted from the individual periodograms are not independent, because segments overlap substantially in our analysis (by design). 
Similarly, powers in averaged periodograms are not independent both of neighbouring segments as well as phase-matched segments, since each power is averaged at least twice with either neighbour (or more times in the case of constructing averaged periodograms from a larger number of rotational cycles). 
In each case, we apply the same analysis to our simulations, and thus construct averaged periodograms in the same way for each of our $N_\mathrm{sim}$ simulations. Consequently, we can construct simple p-values for the significance of a signal in the averaged periodograms. 
 

\section{Results}
\label{sec:results}

\subsection{RXTE}
\label{sec:rxte_results}

We repeated the analysis from \citealt{strohmayer06}, paying special attention to the overall number of cycles during which the signal was present, as well as the duration of the presence of the QPO in any individual cycle.
We search individual segments of $3 \, \mathrm{s}$ length, starting every $0.5 \, \mathrm{s}$, such that consecutive segments overlap by $2.5 \, \mathrm{s}$. We extracted the power at $625 \, \mathrm{Hz}$, then ran 30000 simulations as described in \ref{sec:analysis} and compared the powers at $625 \, \mathrm{Hz}$ for each segment as well as phase-averaged segments to the powers extracted in the same way from the simulations.

There is a strong QPO signal at $625$ Hz in a cycle starting at $239\, \mathrm{s}$ after the trigger, as reported in \citet{strohmayer06}. In order to answer the question whether this signal is present over all nine cycles, for which results are reported in \citet{strohmayer05}, or whether it is only present in a single cycle, we constructed phase-averaged periodograms averaging up to nine periodograms, and constructed p-values as described in Section \ref{sec:analysis}. Figure \ref{fig:rxte_pvals} presents the results: the QPO is significant when averaging together up to $9$ cycles to $p < 0.02$. However, we also note that the signal is most significant when averaging the cycle with the strongest signal and the immediately previous cycle together. This indicates that the QPO is likely to be present at least in these two cycles. When averaging more than two cycles, the significance drops (the p-value rises, i.e. the null hypothesis, pure noise and no QPO, becomes more likely), indicating that with every additional cycle, we are averaging in more noise rather than more signal. We conclude that the signal is most likely present in two cycles at most.
We also note that we fail to reproduce the marginal detection of a QPO at the same frequency six cycles before the cycle with the strongest incarnation of the $625\, \mathrm{Hz}$ QPO. 

Given that the QPO is likely confined to two cycles, it is reasonable to ask whether the QPO extends over both cycles, i.e. is present for the entire $\sim 14\, \mathrm{s}$, or is being re-excited. To test this, we attempted to constrain the width of the QPO in the strongest cycle, in two ways. First, we redid the analysis described above, but with time segments starting exactly one cycle of the $625 \, \mathrm{Hz}$ QPO apart, i.e. every $1/625 = 1.6 \times 10^{-3} \, \mathrm{s}$ apart, effectively providing a finely resolved sliding window over the cycle where the QPO is strongest. If one then plots the strength of the signal with time, one can track the strength of the QPO over the course of the star's rotational cycle. As more signal is included in a given segment, the power will rise, until the entire signal is included. Similarly, as the sliding window moves out of the time frame where the QPO is located, less and less signal is included, and the power drops. We show the resulting plot in Figure \ref{fig:rxte_combined}. Similarly to Figure 3 in \citet{strohmayer06}, the QPO seems to be present only on the declining edge of the second pulse, and only for a short period of time. We then injected an artificial sinusoidal signal into a single cycle in 50 simulations, in the same part of the rotational cycle as the real QPO, for a duration of $0.25\, \mathrm{s}$ and a fractional rms amplitude of $0.3$. The sinusoid is on top of a fast-rise exponential decay profile, to simulate the effect of a suddenly appearing, then decaying signal. We show a single realisation in the lower panel of Figure \ref{fig:rxte_combined}. In the upper panel, we show an image of the simulations, with data over plotted. The data traces the simulations as well, indicating that the observed QPO can be well modelled with a short, decaying, but high-amplitude QPO signal. 


\subsection{RHESSI}
\label{sec:rhessi_results}

\citealt{watts06} searched segments of $t_{\mathrm{set}} = 2.27$ seconds length, i.e. $1/3$ of the neutron star's rotational cycle, over a range of 19 successive cycles, starting $\sim 80 \, \mathrm{s}$ after the onset of the giant flare. We varied the length of the segments between $0.5 \, \mathrm{s}$ and $2.0 \, \mathrm{s}$ in order to be sensitive to shorter signals, which may be buried in noise when taking the periodogram over too long a segment. We subdivided each cycle into $N_\mathrm{s} = 30$ segments, such that they start every $d t_\mathrm{set}= 7.6022/30 = 0.2534 \, \mathrm{s}$ apart, and overlap for $\delta t_\mathrm{set} - 0.2534 \, \mathrm{s}$.


TO ADD:
- periodogram of individual signal in strongest cycle
- show a savg example plot?
- pvalues
- simulation with similar p-values


\section{Discussion}
\label{sec:discussion}


\end{document}
